Earth_microbiome_MS

### 2019-07-16 Earth_microbiome_MS.Rmd
updated

#Check sparsity of physeq data
kable(head(tax_table(githubdata_phyloseq)))
tax_table(githubdata_phyloseq)

@@

#Filter out low-prevalance phyla
  #1. Define phyla to filter
filterPhyla = c("p__Caldiserica", "p__Deferribacteres", "p__LD1", "p__MVP-21", "p__OctSpA1-106", "p__TPD-58", "p__SC4", "p__VHS-B3-43", "p__WS1" )
  #2. Filter entries
github_filtered = subset_taxa(githubdata_phyloseq, !Phylum %in% filterPhyla)
github_filtered
  #3. Create data frame in R Markdown for easier comparison (see prevdf)
prevdf_filtered = apply(X = otu_table(github_filtered),
               MARGIN = ifelse(taxa_are_rows(github_filtered), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

@@

#READ ME - Problems encountered

@@

Some of the phyla in the table are named following a differenct convention. If this file is used for the data this will have to be kept in mind. It is possible that other files for EMP data also share this problem.

>>>>

#Check sparsity of physeq data
#kable(head(tax_table(githubdata_phyloseq)))
#tax_table(githubdata_phyloseq)

@@

#Filter out low-prevalance phyla
  #1. Define phyla to filter
filterPhyla = c("p__Caldiserica", "p__Deferribacteres", "p__LD1", "p__MVP-21", "p__OctSpA1-106", "p__TPD-58", "p__SC4", "p__VHS-B3-43", "p__WS1" )
  #2. Filter entries
github_filtered = subset_taxa(githubdata_phyloseq, !Phylum %in% filterPhyla)
github_filtered
  #3. Create data frame in R Markdown to display filtered prevalence to allow for an easier comparison (see prevdf)
prevdf_filtered = apply(X = otu_table(github_filtered),
               MARGIN = ifelse(taxa_are_rows(github_filtered), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

@@

#READ ME

####Problems encountered

@@

Some of the phyla in the table are named following what appears to be a different convention. However, further research indicates thas these are uncharacterised phyla (likely with no cultured representatives) present in the Silva and Greengenes databases. If this file is used for the data this will have to be kept in mind.

####Notes

I believe taxonomic filtering should not be arbitrary. This can be done by manually choosing phyla and writing them into the function but can likely be done mathematically by choosing a minimum prevalence (total or average)
Prevalence filtering can be used as an alternative (should be consistent with the gut microbiome filtering?)

AC1 (doi: https://dx.doi.org/10.1073%2Fpnas.1015676108)
